A Systematic Approach to Model Predictive Controller Constraint Handling: Rigorous Geometric Methods

### 2011-04-28 article.tex
- Minor fixes - Subfigure numbers [embedded]

\address{Department of Chemical Engineering, University of Pretoria, Pretoria 0001, South Africa}

@@

%% ==== BODY ====
\section{Introduction}\label{sec:intro}
Model predictive controllers use process models to predict the effect of the inputs of a system on its outputs.
The ability of some MPC algorithms to impose constraints on the outputs or inputs is their biggest selling point.
These constraints are also interdependent via the process model.

@@

Since the process model (gain-matrix) operates on and outputs deviation variables, a steady-state offset is added to produce absolute values.
This offset is calculated from the center of the input space and corresponds to the nominal point at which the process model is linearised.

@@

\citet{vinsonartoi} proceed to define the generalised Output Operability Index (OOI) as shown in equation~\ref{eq:oi}. 
From this definition it is clear that the controlability of a process decreases if the DOS does not cover the whole AOS. 
Servo- and regulatory operability indices are also defined in the text, but are not shown here.
\begin{equation}
  \label{eq:oi}
     OOI \triangleq \frac{\mu(AOS\cap DOS)}{\mu(DOS)}
\end{equation}
where $\mu$ represents a function to calculate the hypervolume of a space.

@@

\subsection{Constraint types}\label{sec:contypes}
When constraints are considered, and specifically changes to them, the types of constraints present are important.
This is due to the direction of change allowed by each constraint type.
The following set of constraint types (based on the sources of constraints) are proposed to disambiguate the specification of constraints.
\begin{description}
  \item [Physical constraints] represent the physical limits of the system.
  Examples are tank levels, valve saturation limits and maximum heat duties.
  Allowed changes to these constraints are usually one-sided or to the inside of a range.
  \item [Quality constraints] are concerned with product quality, these constraints are typically determined by external requirements or standards.
  Strictly speaking, changes to these constraints are allowed in any direction.
  \item [Operational constraints] are the innermost constraint set used by the MPC controller.
  These constraints are typically changed using the operator interfaces of MPCs and are contained within their respective variables' physical and quality constraints.
  \item [Optimisation constraints] are those constraints that are not critical for the control of the plant, but rather the performance and efficiency of the plant as determined by external cost functions.
  Typically these external cost functions will represent a cost benefit, e.g. energy savings or product give-away reduction.
\end{description}

@@

\section{2$\times$2 Distillation Column Example}\label{sec:dist}
\subsection{Process Description}
The example case study is a 10-plate distillation column from the Process Modelling and Control laboratory of the University of Pretoria.
The column is run as a closed system with bottoms and distillate mixed and fed back into the column.
Figure~\ref{fig:columnpfd} shows a simplified process flow diagram of the column.
\begin{figure}[htbp]
  \centering
    %\includegraphics[width=8cm]{columnpfd.pdf}
    \scalebox{1}{\input{columnpfd}}
  \caption[Laboratory distillation column photo and flow diagram]{Laboratory distillation column flow diagram.}
  \label{fig:columnpfd}
\end{figure}

The process model has been reduced to a $2\times2$ matrix.
Reflux flowrate ($R$, expressed as the milliamp signal to the control valve) and the setpoint of plate 10's temperature ($T_{10~sp}$) are the inputs, with the temperatures of plate 1 and 8 ($T_1\text{ \& }T_{8}$) as the outputs.
The temperature controller between $T_{10~sp}$ and $Q$ is part of the column's baselayer controls and was added as a safety measure, should the advanced control layer fail.

@@

This set of constraints need to be reformatted to be compatible with commercial MPC packages, as they only accept high and low limits on variables.
An unmeasured variable needs to be added to the process model to implement the diagonal constraint ($0.88T_8-0.47T_1\leq 36.56$~\textcelsius).
Defining the new output $y_1$ and augmenting the process model (equation~\ref{eq:columnmodel}) as per equation~\ref{eq:linconoutput} results in the new process model (equation~\ref{eq:columnnewmodel}) shown below. With the new process model defined thus, the diagonal constraint to be expressed simply as {$y_1~\leq~36.56$~\textcelsius}.
\begin{equation}
  \label{eq:columnnewmodel}
  G_{new}= \bpm -0.0575 & 0.96 \\       % T1
                  -0.146  & 0.518 \\      % T8
                   0.0956 & 1.30 \\ \epm  %y1
\end{equation}

@@

\section{Conclusions}\label{sec:conclusions}
From the proposed method of systematic constraint handling and the illustrative example, the conclusions can be drawn:
\begin{itemize}
  \item Checking constraints via the process model can identify constraints which are not attainable and should be revised.
  \item The constraints describing the AOS and DOS intersection represent the largest operating region for the given input and output constraints.
By adding unmeasured variables to the process model, these linear constraints can be reformatted and directly applied in the high and low limit structure of commercial MPC interfaces.
  \item When constraint changes are made (during operation or design), checking changes to constraints via the process model allows for a measure of clamping (or relaxing) to be determined.
This highlights the effect of changing constraints and can guide design or operation decisions.
\end{itemize}

>>>>

\address{Department of Chemical Engineering, University of Pretoria, Pretoria, 0001, South Africa}

@@

%% ==== BODY ====
\section{Introduction}\label{sec:intro}
Model predictive controllers use process models to predict the effect of the inputs of a system on its outputs.
The ability of some MPC algorithms to impose constraints on the outputs or inputs is one of their biggest selling points.
These constraints are also interdependent via the process model.

@@

Since the process model (gain-matrix) operates on, and outputs, deviation variables, a steady-state offset is added to produce absolute values.
This offset is calculated from the center of the input space and corresponds to the nominal point at which the process model is linearised.

@@

\citet{vinsonartoi} proceed to define the generalised Output Operability Index (OOI) as shown in equation~\ref{eq:oi}. 
\begin{equation}
  \label{eq:oi}
     OOI \triangleq \frac{\mu(AOS\cap DOS)}{\mu(DOS)}
\end{equation}
where $\mu$ represents a function to calculate the hypervolume of a space.

From this definition it is clear that the controlability of a process decreases if the DOS does not cover the whole AOS. 
Servo- and regulatory operability indices are also defined in the text, but are not shown here.

@@

\subsection{Constraint types}\label{sec:contypes}
When constraints are considered, and specifically changes to them, the types of constraints present are important.
This is due to the direction of change allowed for each constraint type.
The following set of constraint types (based on the sources of constraints) are proposed to disambiguate the specification of constraints.
\begin{description}
  \item [Physical constraints] represent the physical limits of the system.
  Examples are tank levels, valve saturation limits and maximum heat duties.
  Allowed changes to these constraints are usually one-sided or to the inside of a range.
  \item [Quality constraints] are concerned with product quality, these constraints are typically determined by external requirements or standards.
  Strictly speaking, changes to these constraints are allowed in any direction.
  \item [Operational constraints] are the innermost constraint set used by the MPC controller.
  These constraints are typically changed using the operator interfaces of MPCs and are contained within their respective variables' physical and quality constraints.
  \item [Optimisation constraints] are those constraints that are not critical for the control of the plant, but rather the performance and efficiency of the plant as determined by external cost functions.
  Typically these external cost functions will represent a cost benefit, e.g. energy savings or product give-away reduction.
\end{description}

@@

\section{2$\times$2 Distillation Column Example}\label{sec:dist}
The method proposed in the preceding section is now illustrated for a practical system.
Even though the method applies to higher order systems, a $2 \times 2$ system is considered to ease graphical representation.
\subsection{Process Description}
The example case study is a 10-plate distillation column from the Process Modelling and Control laboratory of the University of Pretoria.
The column is run as a closed system with bottoms and distillate mixed and fed back into the column.
Figure~\ref{fig:columnpfd} shows a simplified process flow diagram of the column.
\begin{figure}[htbp]
  \centering
    %\includegraphics[width=8cm]{columnpfd.pdf}
    \scalebox{1}{\input{columnpfd}}
  \caption[Laboratory distillation column photo and flow diagram]{Laboratory distillation column flow diagram.}
  \label{fig:columnpfd}
\end{figure}

The process model has been reduced to a $2\times2$ matrix.
Reflux flowrate ($R$, expressed as the milliamp signal to the control valve) and the setpoint of plate 10's temperature ($T_{10~sp}$) are the inputs, with the temperatures of plate 1 and 8 ($T_1\text{ and }T_{8}$) as the outputs.
The temperature controller between $T_{10~sp}$ and $Q$ is part of the column's baselayer controls and was added as a safety measure, should the advanced control layer fail.

@@

This set of constraints need to be reformatted to be compatible with commercial MPC packages, as they only accept high and low limits on variables.
An unmeasured variable is added to the process model to implement the diagonal constraint ($0.88T_8-0.47T_1\leq 36.56$~\textcelsius).
Defining the new output $y_1$ and augmenting the process model (equation~\ref{eq:columnmodel}) as per equation~\ref{eq:linconoutput} results in the new process model (equation~\ref{eq:columnnewmodel}) shown below. With the new process model defined thus, the diagonal constraint can be expressed as {$y_1~\leq~36.56$~\textcelsius}.
\begin{equation}
  \label{eq:columnnewmodel}
  G_{new}= \bpm -0.0575 & 0.96 \\       % T1
                  -0.146  & 0.518 \\      % T8
                   0.0956 & 1.30 \\ \epm  %y1
\end{equation}

@@

\section{Conclusions}\label{sec:conclusions}
From the proposed method of systematic constraint handling and the illustrative example, these conclusions can be drawn:
\begin{itemize}
  \item Checking constraints via the process model can identify constraints which are not attainable and should be revised.
  \item The constraints describing the AOS and DOS intersection represent the largest operating region for the given input and output constraints.
By adding unmeasured variables to the process model, these linear constraints can be reformatted and directly applied in the high and low limit structure of commercial MPC interfaces.
  \item When constraint changes are made (during operation or design), checking changes to constraints via the process model allows for a measure of clamping (or relaxing) to be determined.
This highlights the effect of changing constraints and can guide design or operation decisions.
\end{itemize}
